---
title: "Data Import; Exploratory Data Analysis (EDA) in R"
output: pdf_document
urlcolor: "blue"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3, 
                      fig.width  = 6,
                      fig.align  = "center")
```

Based on lecture handouts originally written by Dr. David Gerard.

# Learning Objectives

- Import data from CSV's, 
- Working Directories
- Strategies for EDA
- [Data Import Cheat Sheet](https://github.com/rstudio/cheatsheets/blob/master/data-import.pdf).
- [Readr Overview](https://readr.tidyverse.org/).

# Part 1 (Data Import)

# Working Directories

- The working directory is where R will look for and save things by default.

- When you specify to save a figure, save a file, or load some data, it will be
  with respect to the working directory.
  
- You can see where the current working directory is by `getwd()`, or by looking
  at the top of the console in RStudio.

- You can change the working directory by Session > Set Working Directory > 
  Choose Directory. Or by CONTROL + SHIFT + H. Or you can use the `setwd()` 
  command.
  
- A shortcut is to set the working directory to your source file location with
  Session > Set Working Directory > To Source File Location.
  
- When you read and write files/figures, you can then specify the path
  from the position of the working directory.

- Suppose we want to save the following figure:  
    ```{r}
    suppressPackageStartupMessages(library(tidyverse))
    data("mpg")
    pl <- ggplot(mpg, aes(x = hwy, y = cty)) +
      geom_point()
    ```

- To save `pl` in the current folder, we would use:
    ```{r, eval = FALSE}
    ggsave(filename = "./my_saved_plot.pdf", plot = pl)
    ```
    
- The "`.`" means "the current folder".
    
- To save `pl` in the folder one level up we would use:
    ```{r, eval = FALSE}
    ggsave(filename = "../my_saved_plot.pdf", plot = pl)
    ```

- The "`..`" means "go one level up".

- If we are in the analysis folder, and we want to save `pl` in the output 
  folder, we would use:
  
    ```{r, eval = FALSE}
      ggsave(filename = "../output/my_saved_plot.pdf", plot = pl)
    ```
    
- If we have a subfolder called "fig" within out current folder. We could
  save `pl` in "fig" with
    ```{r, eval = FALSE}
      ggsave(filename = "./fig/my_saved_plot.pdf", plot = pl)
    ```
    
- **NEVER USE ABSOLUTE PATHS**. For example, you should never start the path
  from "C" if you use Windows. This makes your code non-transferable to other
  users.
  
# readr

- To read a CSV (comma-separated values) file into R, use the `read_csv()` 
  function from the readr package.
  
    ```{r}
    suppressPackageStartupMessages(library(tidyverse))
    heights <- read_csv(file = "./heights.csv")
    ```
    

- Use `read_tsv()` if columns are separated by tabs.

- Use `read_csv2()` if columns are separated by semicolons.

- Other file formats are listed in [RDS](https://r4ds.had.co.nz/).

- First export the Excel spreadsheet as a CSV. Then read the CSV file into R.

- You are using colors to represent meaningful information in Excel? Don't.

    - Edit the data so that the information is encoded by a new variable.
    
- If you don't know the format ahead of time, use `read_lines()` to print
  the first few lines.
  
    ```{r}
    read_lines(file = "./heights.csv", n_max = 10)
    ```
  
    
# Special Considerations

- **Always check your data immediately after importing it**.

    - Check that the types are correct for each of the variables.
    - Check that the missing data were coded correctly.
    - Later on, when you notice something weird, consider that this might have
      resulted because of a problem during data import.
    
    ```{r}
    hate_crimes <- read_csv(file = "./hate_crimes2.csv")
    
    summarize_all(hate_crimes, class)
    
    summarize_all(hate_crimes,funs(sum(is.na(.))))#sum all the NA's under each variable
    
    head(hate_crimes)
    ```
    

- Sometimes the files code missing data other than `NA`. For example, it's common
  to use periods `.`, or in some genomic settings they use `-9` as missing.
  
- R won't know how to handle this without you telling it, so you'll have to 
  know what the missing data encoding is and specify it with the `na` argument
  in `read_csv()`.
  
- readr will try to guess the type for each column (double, integer, character,
  logic, etc). Sometimes it guesses wrong. If it seems to be guessing wrong,
  use the `col_types` to explicitly specify the column types.
  
- Sometimes there are comments at the start of a data file. You can skip the 
  first few lines before starting to read data with the `skip` argument.
  
- If the comments begin with a special character, you can use the `comment` 
  argument.
  
# Data Export

- You can write comma-separated and tab-separated files using `write_csv()`,
  `write_csv2()`, and `write_tsv()`.
  
- The defaults are usually fine.

# Reading/Writing R Objects

- You can save and reload arbitrary R objects (data frames, matrices, lists, 
  vectors) using `readRDS()` and `saveRDS()`.

# Part 2 (Exploratory Data Analysis (EDA) in R)

  We will use ggplot2 which is a R package dedicated to data visualization. 

# General Strategies

- Plot the distribution of every variable.
- Plot the bivariate distribution of every pair of variables (to find which
  variables are associated).
- Color code by variables to try and see if relationships can be explained.
- Calculate lots of summary statistics.
- Look at missingness.
- Look at outliers.
- EDA is about **curiosity**. Ask *many* questions, use *many* plots, 
  investigate *many* aspects of your data. This will let you hone in on 
  the few *interesting* questions you want to pursue deeper.
  
```{r, message=FALSE}
library(tidyverse)
data("diamonds")
```

  
# Distribution of Every Variable:

- Quantitative: Use a histogram.
    
    - Look for modality. Indicates multiple groups of units. What can explain 
      the modes? Can any of the other variables explain the modes?
    - Are certain values more likely than other values?
    - Look for skew. 
    - `geom_histogram()`
    - [Mean](https://en.wikipedia.org/wiki/Mean), 
      [median](https://en.wikipedia.org/wiki/Median), 
      [standard deviation](https://en.wikipedia.org/wiki/Standard_deviation),
      [five number summary](https://en.wikipedia.org/wiki/Five-number_summary).

    ```{r}
    ggplot(data = diamonds, mapping = aes(x = carat)) +
      geom_histogram(bins = 500)
    
    fivenum(diamonds$carat)
    mean(diamonds$carat)
    sd(diamonds$carat)
    ```

- Categorical: Use a bar chart. Or just a table of *proportions* (`table()` then
  `prop.table()`).

    - Absolute counts are sometimes interesting, but usually you want to 
      look at the proportion of observations in each category.
    - Is there a natural ordering of the categories (bad, medium, good)?
    - Why are some categories more represented than others?
    - `geom_bar()`, `geom_col()`
    - Proportion of observations within each group.
    
    ```{r}
    ggplot(diamonds, aes(x = color, y = )) +
      geom_bar(aes(y = count / sum(..count..))) +
      ylab("Proportion")
    
    table(diamonds$color)
    prop.table(table(diamonds$color))
    ```

# Bivariate Distribution of Every Pair of Variables

- Quantitative vs Quantitative: Use a scatterplot
    
    - Is the relationship linear? Quadratic? Exponential?
    - Logging is useful tool to make some associations linear. 
      If the relationship is (i) monotonic and (ii) curved, then try logging 
      the x-variable *if the x-variable is all positive*.
      If it is also (iii) more variable at larger y-values, then try logging
      the y-variable *instead* of the x-variable *if the y-variable is all 
      positive*. Try logging both if you still
      see curvature *if both variables are all positive*.
    - Ask if an observed association can be explained by another variable?
    - Correlation coefficient (only appropriate if association is linear). 
    - Kendall's tau (always appropriate).
    
    ```{r}
    ggplot(diamonds, aes(x = carat, y = price)) +
      geom_point() +
      scale_y_log10() +
      scale_x_log10()
    
    cor(diamonds$carat, diamonds$price)
    ## cor(diamonds$carat, diamonds$price, method = "kendall")
    ```
    
- Categorical vs Quantitative: Use a boxplot

    - For which levels of the categorical variable is the quantitative variable
      higher or lower?
    - For which levels is the quantitative variable more spread out?
    - Aggregated means, medians, standard deviations, quantiles
    
    ```{r}
    ggplot(diamonds, aes(x = color, y = price)) +
      geom_boxplot() +
      scale_y_log10()
    
    diamonds %>%
      mutate(logprice = log(price)) %>%
      group_by(color) %>%
      summarize(mean   = mean(logprice), 
                sd     = sd(logprice), 
                median = median(logprice),
                Q1     = quantile(logprice, 0.25), 
                Q3     = quantile(logprice, 0.75))
    ```
    

- Categorical vs Categorical: Use a mosaic plot or a count plot

    - For which pairs of values of the categorical variables are there the most number of 
      units?

    ```{r}
    ## Only gives you the bivariate distribution
    ggplot(diamonds, aes(x = cut, y = color)) +
      geom_count()
    
    ## Gives you the conditional distributions of color given cut
    ggplot(diamonds, aes(x = cut, fill = color)) +
      geom_bar(position = "fill")
    
    ## Gives you the conditional distributions of cut given color
    ggplot(diamonds, aes(x = color, fill = cut)) +
      geom_bar(position = "fill")
    ```














