---
title: "STAT 690 DATADownload"
author: "Wei-Chen Lu"
date: "2021/6/21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Dowmload proper library for loading datasets and cleaning the data for better visualization.
```{r}
#install.packages("tidyverse")
#install.packages("Readr")
```

# Adult Dataset
-Extraction was done by Barry Becker from the 1994 Census database. Prediction task is to determine whether a person makes over 50K a year.

# Method 1: Download the csv from file directly
-To be more specific, students can download the csv file from Canvas directly to them own pc's desktop and then open a Rmd file and store it at desktop as well. After doing this, students can apply this code to open the data set directly.

```{r}
read.csv("adult1.csv")->adult1
head(adult1)
```

#Method 2: Download the raw dataset from the open data website and modify for better use
-This adult data set can be download from open data website: https://urldefense.com/v3/__https://archive.ics.uci.edu/ml/datasets/adult__;!!IaT_gp1N!hnmhRF4qpBA4ZedEiEcPs7I_HmX8YqDRz6Z_ktl0Uw1HApxDGW_8Vd7BS7MICSWa$ .
-We can download the data through the hyper link and then input column names for better visualization.

```{r}
adult1<-read.csv(url("https://urldefense.com/v3/__https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data__;!!IaT_gp1N!hnmhRF4qpBA4ZedEiEcPs7I_HmX8YqDRz6Z_ktl0Uw1HApxDGW_8Vd7BSyZ9MjYS$ "), header = F)
#this dataset's column names are not defined and we can put the column names according to the website's description.
colnames(adult1)<-c("age","workclass",
                   "fnlwgt","eduaction",
                   "education_num",
                   "marital_status",
                   "occupation",
                   "relationship",
                   "race","sex",
                   "capital_gain",
                   "capital_loss",
                   "hours_per_week",
                   "native_country",
                   "income")  #fct variable
head(adult1)
```

