---
title: "STAT 690 DATADownload"
author: "Wei-Chen Lu"
date: "2021/6/21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Dowmload proper library for loading datasets and cleaning the data for better visualization.
```{r}
#install.packages("tidyverse")
#install.packages("Readr")
```


# Depression Dataset
-Depression data set is from A study of 1,594 high school students has been conducted in order to find socioeconomic and family factors that may be associated with stress and depression.

# Method 1: Download the csv from file directly
-To be more specific, students can download the csv file from Canvas directly to them own pc's desktop and then open a Rmd file and store it at desktop as well. After doing this, students can apply this code to open the data set directly.
```{r}
read.csv("depression_data.csv")->dp
head(dp)
```

#Method 2: Download the dataset from website hyperlink:
-This depression dataset can be found on several open data websites, and here we prepare to download this dataset through a super link from the AU's course website which made by Dr.Baron, a statistic machine learning specialist working as Professor in AU for over 15 years.  

"http://fs2.american.edu/baron/www/627/" This is a website allow people who have strong interests to learn and to work with machine learning in R.
```{r}
read.csv(url("http://fs2.american.edu/~baron/627/R/depression_data.csv%22))-> dp
head(dp)
```
